---
title: "ggtree Foundations: Topology, Layout, and Annotation"
output: html_document
---

```{r setup, message=FALSE}
library(phyloheatmapviz)
tree <- ape::read.tree("../data/pathogen_tree.nwk")
metadata <- read.csv("../data/tip_metadata.csv")
```

## Validate before joining

```{r validation}
validate_tip_alignment(tree, metadata)
```

Tree tip labels are keys. A row-number join can silently attach host, lineage,
or resistance information to the wrong isolate.

## Rectangular and circular layouts

```{r trees}
rectangular_phylogeny(tree, metadata)
circular_phylogeny(tree, metadata)
```

The same topology can be drawn in different layouts. Layout is presentation;
it does not change evolutionary distance or clade membership.

## Support and aligned genomic traits

```{r support}
clade_support_tree(tree, support_threshold = 80)
```

```{r pangenome}
pangenome <- read.csv("../data/pangenome.csv", row.names = 1, check.names = FALSE)
aligned_pangenome_tree(tree, as.matrix(pangenome))
```

The pangenome matrix is reordered to `tree$tip.label` only after exact set
equality is established.

