---
title: "ComplexHeatmap Workflows for Omics"
output: html_document
---

```{r setup, message=FALSE}
library(phyloheatmapviz)
read_matrix <- function(path) {
  x <- read.csv(path, check.names = FALSE)
  ids <- x[[1]]; x <- as.matrix(x[-1]); rownames(x) <- ids
  storage.mode(x) <- "numeric"; x
}
samples <- read.csv("../data/sample_metadata.csv")
expression <- read_matrix("../data/expression_zscores.csv")
```

## Annotation-aware expression heatmap

```{r expression}
ComplexHeatmap::draw(annotated_expression_heatmap(expression, samples))
```

The function requires metadata in exactly the same order as matrix columns.
Condition defines the split; batch remains visible as an annotation rather than
being mistaken for a biological cluster.

## Multi-omics alignment

```{r multiomics}
pathway <- read_matrix("../data/pathway_activity.csv")
ComplexHeatmap::draw(multiomics_heatmap(expression, pathway, samples))
```

Concatenated heatmaps share sample order. A mismatch is an error rather than a
plausible-looking but incorrect figure.

## Sample-similarity review

```{r correlation}
ComplexHeatmap::draw(correlation_heatmap(expression, samples))
```

Correlation structure can identify potential batch patterns or outliers, but
the result still depends on filtering, normalization, and feature selection.

